Baseline centering problem

### 2019-01-15 paper/tangotd.tex

Normalization of $Z$. Since $Z=\sum \gamma^t P^t$, we have that
$(1-\gamma)Z$ is a stochastic matrix (thanks Yann). This might help with
normalization/detecting divergence. This translates as $\E_{s_2\sim \rho}
z(s_1,s_2)=\frac{1}{1-\gamma}$ which is testable/enforceable. However, I
fear it might make the fixed point $z(s_1,s_2)=\frac{1}{1-\gamma}$
attractive instead of repulsive, and it might also perturb initialization
(initializing $Z=\Id$ does not satisfy this normalization, so enforcing
it too early is basically like initializing to
$Z=\frac{1}{1-\gamma}\Id$ which empirically diverges).

@@

It's a bit strange to train the policy for goal $s_2$ via policy gradient,
instead of just selecting the next state to maximize
$z(s,s_2;s_2)$ where $z(s,s_2;s_3)$ is the $z$ function associated with
the policy that has goal $s_3$. Formally, computing the latter is
feasible: one has to follow an exploration policy that has a non-constant
goal,so at each step we observe a transition of the policy with some goal
$s_3$; and hope that Newton--Bellman for computing $z(s,s_2;s_3)$
generalizes well with respect to $z_3$.

@@

By multiplying \eqref{eq:optleftbellman} by $V$ on the left, one finds the
Newton--Bellman equation for the optimal function $V$ seen as a matrix:
\begin{equation}
V^2_{s_1s_2}=V_{s_1s_2}+\gamma \sum_{s_3} V_{s_1s_3}\max_a
\E_{s_4|s_3,a} V_{s_4s_2}
\end{equation}
and representing $V$ by its density wrt $\rho$ namely
$V_{s_1s_2}=z^\ast(s_1,s_2)\rho(s_2)$ this suggests the Newton method
\begin{multline}
z^\ast(s_1,s_2)\gets 2z^\ast(s_1,s_2)-\E_{s_3\sim \rho}\,
z^\ast(s_1,s_3)z^\ast(s_3,s_2)\\
+\gamma
\E_{s_3\sim \rho} \,z^\ast(s_1,s_3)\max_a \E_{s_4|s_3,a} z^\ast(s_4,s_2)
\end{multline}
satisfied by the optimal $z$ function. This is not quite an instance of
the Newton method for matrix inversion anymore (because the operator
$\Id-\gamma \max_a \E_{\cdot|a}$ is not linear), so convergence
is unclear.

>>>>

Normalization of $Z$. Since $Z=\sum \gamma^t P^t$, we have that
$(1-\gamma)Z$ is a stochastic matrix (thanks Yann). This might help with
normalization/detecting divergence. This translates as $\E_{s_2\sim \rho}
z(s_1,s_2)=\frac{1}{1-\gamma}$ which is testable/enforceable. However, I
fear it might make the fixed point $z(s_1,s_2)=\frac{1}{1-\gamma}$
attractive instead of repulsive, and it might also perturb initialization
(initializing $Z=\Id$ does not satisfy this normalization, so enforcing
it too early is basically like initializing to
$Z=\frac{1}{1-\gamma}\Id$ which empirically diverges). Still, one might
use whether $\E_{s_2\sim \rho}z(s_1,s_2)$ is significantly above
$\frac1{1-\gamma}$ as a divergence criterion (same with negative values).

@@

It's a bit strange to train the policy for goal $s_2$ via policy gradient
on the policy with parameter $\omega(s_2)$, instead of just selecting the
policy parameter $\omega$ to maximize $z(s,s_2;\omega)$ where
$z(s,s_2;\omega)$ is the $z$ function associated with the policy that has
goal $\omega$.  This is why we propose to train the function $s_2\mapsto
\omega$ so that $\omega(s_2)$ is indeed the parameter that maximizes
$z(s,s_2;\omega)$.
To compute $z(s,s_2;\omega)$ one has to follow an
exploration policy that has a non-constant goal, so at each step we
observe a transition of the policy with some goal $\omega$; and hope that
Newton--Bellman for computing $z(s,s_2;\omega)$ generalizes well with
respect to $\omega$.

@@

By multiplying \eqref{eq:optleftbellman} by $V$ on the left, one finds the
Newton--Bellman equation for the optimal function $V$ seen as a matrix:
\begin{equation}
V^2_{s_1s_2}=V_{s_1s_2}+\gamma \sum_{s_3} V_{s_1s_3}\max_a
\E_{s_4|s_3,a} V_{s_4s_2}
\end{equation}
and representing $V$ by its density wrt $\rho$ namely
$V_{s_1s_2}=z^\ast(s_1,s_2)\rho(s_2)$ this suggests the Newton-like method
\begin{multline}
z^\ast(s_1,s_2)\gets 2z^\ast(s_1,s_2)-\E_{s_3\sim \rho}\,
z^\ast(s_1,s_3)z^\ast(s_3,s_2)\\
+\gamma
\E_{s_3\sim \rho} \,z^\ast(s_1,s_3)\max_a \E_{s_4|s_3,a} z^\ast(s_4,s_2)
\end{multline}
satisfied by the optimal $z$ function. This is not quite an instance of
the Newton method for matrix inversion anymore (because the operator
$\Id-\gamma \max_a \E_{\cdot|a}$ is not linear), so convergence
is unclear.
